


```{r setup, include=FALSE}
knitr::opts_chunk$set(
    # 在 markdown 文档显示代码块
    echo = TRUE, 
    
    # 不显示警告信息
    warning = FALSE, 
    
    # 在新块中写入代码的输出结果
    collapse = TRUE, 
    
    # 所有代码运行完成后再输出
    results = 'hold', 
    
    # 输出代码前用#号注释
    comment='#'
)
```


# Quick start

## About

+ Although with bright prospects in Pan-disease analysis, [GSClassifier](https://github.com/huangwb8/GSClassifier) was primarily developed for clinic-friendly immune subtypes of gastric cancer (GC). Currently, only `PAD` subtypes and `PADi` for GC were supported. We would try to support more cancer types in the future as possible. More details in [Plans in the future](https://github.com/huangwb8/GSClassifier/wiki/Plans-in-the-future) section.
+ Gibbs\' `PanCancer immune subtypes` based on five gene signatures (485 genes) could also be called in `GSClassifier`, with a pre-trained model from the [ImmuneSubtypeClassifier](https://github.com/CRI-iAtlas/ImmuneSubtypeClassifier) package. If you use their jobs, please cite [these papers](https://github.com/huangwb8/GSClassifier/wiki/Introduction#Reference).
+ Particularly, all normal tissues should be eliminated before subtype identificaiton for cancer research.

## Package

```{r}
# Install "devtools" package
if (!requireNamespace("devtools", quietly = TRUE))
  install.packages("devtools")

# Install dependencies
if (!requireNamespace("luckyBase", quietly = TRUE))
  devtools::install_github("huangwb8/luckyBase")

# Install the "GSClassifier" package
if (!requireNamespace("GSClassifier", quietly = TRUE))
  devtools::install_github("huangwb8/GSClassifier")

# Load needed packages
library(GSClassifier)
```

## Data

To lower the learning cost of `GSClassifier`, we provide some test data: 

```{r}
testData <- readRDS(system.file("extdata", "testData.rds", package = "GSClassifier"))
```


Explore the `testData`:

```{r}
names(testData)
```

## PAD

### Preparation of the test data

Load phenotype data:

```{r}
design <- testData$PanSTAD_phenotype_part
table(design$Dataset)
```

Load target sample IDs in `GSE54129` cohort:

```{r}
target_ID <- design$ID[design$Dataset %in% 'GSE54129']
expr <- testData$PanSTAD_expr_part[,target_ID]
head(expr[,1:10])

```


### Unsupervised clustering

```{r out.width='60%', fig.align='center'}
res_pad <- PAD(
  expr = expr,
  cluster.method = "ward.D2",
  extra.annot = NULL,
  plot.title = 'PAD test',
  subtype = "PAD.train_20220916",
  verbose = T
)
```


### Of note

+ It's strongly recommended that the gene type of `expr` should be always the same, such as ENSEMBL genes (ENSG00000111640 for GAPDH, for example).

+ `PAD` function is only for datasets with lots of samples for its classification depends on population-based unsupervised clustering. `PAD` is population-dependent and non-personalized.

+ Beta characteristics: You could try random forest classification based on the `randomForest` package or methods in `stats::hclust`. 

## PADi

+ In `GSClassifier`, `PADi` is a pre-trained out-of-the-box model for GC personalized PAD subtypes calling.

+ During the subtype calling, the gene rank relations based on individuals instead of the relative values across samples would be used. Thus, **you don't have to do batch normalization** even though the data (the `X` input) come from multiple cohorts or platforms.

+ More limitations were discussed in our paper that you had better know.

In this section, we would show how to use `PADi` function series: `PADi`, `callEnsemble`, and `parCallEnsemble` functions.

### Preparation of the test data

```{r}
X <- testData$Kim2018_3
head(X)
```

### Use a specific function called `PADi`

Very simple, just:

```{r warning=FALSE, error=FALSE, message=FALSE, eval=TRUE, results='hide'}
res_padi <- PADi(X = X, verbose = F)
```

Check the result:

```{r}
head(res_padi)
```


Actually, `PADi` is exactly based on a general function called `callEnsemble`.


### Use the `callEnsemble` function

Also simple, just:

```{r warning=FALSE, error=FALSE, message=FALSE, eval=TRUE, results='hide'}
res_padi <- callEnsemble(
    X = X,
    ens = NULL,
    geneAnnotation = NULL,
    geneSet = NULL,
    scaller = NULL,
    geneid = "ensembl",
    subtype = "PAD.train_20220916",
    verbose = F
  )

```

Check the result:

```{r}
head(res_padi)
```


### Parallel strategy for PADi

+ Sometimes, the number of patients for subtype callings could be huge (hundreds or even tens of thousands). Thus, parallel computing (Windows or Linux pass; not tested in Mac or other OS) was also developed in the current version of the `GSClassifier` package.

+ The parameter `numCores` was used to control the No. of CPU for computing (which depends on your CPU capacity).

```
# No run for the tiny test data. With errors.

# Method 1:
res_padi <- PADi(X = X, verbose = F, numCores = 4)

# Method 2: 
res_padi <- parCallEnsemble(
  X = X,
  ens = NULL,
  geneAnnotation = NULL,
  geneSet = NULL,
  scaller = NULL,
  geneids = 'ensembl',
  subtype = 'PAD.train_20220916',
  verbose = T,
  numCores = 4)
```


### Single sample subtype calling

In clinical practice, the single sample subtype calling might be one of the most common scenarios and is also supported by functions of the `PADi` series.

Supposed that there is a GC patient, its information should be:

```{r}
X_ind <- X[,1]; names(X_ind) <- rownames(X)
head(X_ind)
```

Or it can also be another format:

```{r}
X_ind <- as.matrix(X[,1]); rownames(X_ind) <- rownames(X)
head(X_ind)
```


Similar to multiples sample calling, just:

```{r, include=FALSE, warning=FALSE, error=FALSE, message=FALSE, eval=TRUE, results='hide'}
res_padi <- PADi(X = X_ind, verbose = F)
```

check the result:

```{r}
head(res_padi)
```

Similarly, there is alternative choice:

```{r warning=FALSE, error=FALSE, message=FALSE, eval=TRUE, results='hide'}
res_padi <- callEnsemble(
    X = X_ind,
    ens = NULL,
    geneAnnotation = NULL,
    geneSet = NULL,
    scaller = NULL,
    geneid = "ensembl",
    subtype = "PAD.train_20220916",
    verbose = F
  )
```

Check the result:

```{r}
head(res_padi)
```


### Of note

+ In the results of `PADi`, two types of subtypes (`BestCall` and `BestCall_Max`) were integrated. `BestCall` was predicted via an **xgboost** model based on prior knowledge of `PAD` subtypes and the possibility matrix (columns 4 to 7 of four-subtype calling, for example), while `BestCall_Max` was predicted via maximum strategy. Empirically, `BestCall` seemed to be a better choice.

+ `PADi` is individual-dependent and personalized, which means that the result of subtype calling would not be influenced by the data of others.


## Use external models from luckyModel package

In the future, there might be lots of models available as a resource of `GSClassifier`, such as [luckyModel](https://github.com/huangwb8/luckyModel). Here we show how `luckyModel` support `GSClassifier`.

First, install and load `luckyModel`:

```{r}
# Install luckyModel
if (!requireNamespace("luckyModel", quietly = TRUE))
    devtools::install_github("huangwb8/luckyModel")
library(luckyModel)
```

Check projects supported in current `luckyModel`:

```{r}
list_project()
```

Check available models in the project:

```{r}
list_model(project='GSClassifier')
```
Here, `HWB_PAD_v20200110` is a standard name of `PADi`. They are the same. 

Taking `PADi` as an example, we here show how to use an external model from `luckyModel`. First, load a model:

```{R}
model <- lucky_model(project = 'GSClassifier',
                     developer='HWB',
                     model = 'PAD',
                     version = 'v20200110')
```

Then, check the gene id type:

```{r}
model$geneSet
```

The model should use `ensembl` as the value of `geneid` parameter in `callEnsemble` series.

Next, you can use the model like:

```{r warning=FALSE, error=FALSE, message=FALSE, eval=TRUE, results='hide'}
res_padi <- callEnsemble(
  X = X,
  ens = model$ens$Model,
  geneAnnotation = model$geneAnnotation,
  geneSet = model$geneSet,
  scaller = model$scaller$Model,
  geneid = "ensembl",
  subtype = NULL,
  verbose = F
)
```

Or just:

```{r warning=FALSE, error=FALSE, message=FALSE, eval=TRUE, results='hide'}
res_padi <- callEnsemble(
  X,
  ens = NULL,
  geneAnnotation = NULL,
  geneSet = NULL,
  scaller = NULL,
  geneid = "ensembl",
  subtype = model, 
  verbose = F
)

```

They are exactly the same.

Finally, check the result:

```{r}
head(res_padi)
```


## PanCancer immune subtypes

`GSClassifier` could also call the `PanCancer immune subtypes` of Gibbs' [@RN315].

First, see data available in current `GSClassifier`:

```{r}
GSClassifier_Data()
```

Let's use our test data to do this:

```{r}
X <- testData$Kim2018_3
symbol <- convert(rownames(X))
rownames(X) <- symbol
X <- X[!is.na(symbol),]
dim(X)
```

Have a check

```{r}
head(X)
```

PanCan Immune Subtype callings: 

```{r warning=FALSE, error=FALSE, message=FALSE, eval=TRUE, results='hide'}
res_pis <- callEnsemble(
    X = X,
    ens = NULL,
    geneAnnotation = NULL,
    geneSet = NULL,
    scaller = NULL,
    geneid = "symbol",
    subtype = "ImmuneSubtype",
    verbose = F
  )
```

Check the result:

```{r}
head(res_pis)
```


Also, you can try to use `luckyModel`:

```{r}
pci <- lucky_model(
  project = "GSClassifier",
  model = "PanCancerImmuneSubtype",
  developer = "Gibbs",
  version = "v20190731"
)

```


PanCan Immune Subtype callings: 


```{r warning=FALSE, error=FALSE, message=FALSE, eval=TRUE, results='hide'}
res_pis <- callEnsemble(
    X = X,
    ens = NULL,
    geneAnnotation = NULL,
    geneSet = NULL,
    scaller = NULL,
    geneid = "symbol",
    subtype = pci,
    verbose = F
  )
```


Finally, we take a look at the `PanCancer immune subtypes` model:

```{r}
ImmuneSubtype <- readRDS(system.file("extdata", "ImmuneSubtype.rds", package = "GSClassifier"))
names(ImmuneSubtype)
```

Its gene annotation:

```{r}
head(ImmuneSubtype$geneAnnotation)
```

Its gene sets:

```{r}
ImmuneSubtype$geneSet
```
Enjoy `GSClassifier`! 
